A higher order approximate static condensation method for multi-material diffusion problems

### 2018-12-21 asc.tex
Flip the sign of the affine shift for ext fluxes dofs
Eugene's changes

	Multi-phase and multi-material processes are ubiquitous in nature and engineering.
	Mathematical models of such processes are often formulated in terms of systems of partial differential equations with discontinuous coefficients and interface conditions on time-dependent internal boundaries.
	Due to their high complexity, these models   resist analytical treatment, and hence numerical simulations became an indispensable tool for solving the models and gaining insights into underlying phenomena.

	Past decades evidenced an 	explosive growth in developing of efficient computational techniques for problems with evolving interfaces, which involve interaction between many materials or different phases of matter. This includes immersed boundary method \cite{peskin2002immersed}, level-set method \cite{sethian1999level}, and unfitted finite element methods \cite{belytschko2009review,burman2015cutfem}, to name several popular numerical approaches.
	The present study is motivated by the successful applications of the volume-of-fluid (VOF)~\cite{hirt1981volume} and the moment-of-fluid (MOF) methods~\cite{ahn2009adaptive} for the numerical  reconstruction of time-dependent interfaces. Both  methods do not require a mesh aligned with the  interfaces and reconstruct them from the information about the volumes of materials in each cell.
	% implicitly with the help of a characteristic function
	% and construct allow for general polygonal computational meshes not fitted to the position of the interface.
	The superior conservation properties of the  VOF and MOF approaches, however, commonly come at a price of recovering piecewise smooth interface approximation that is  discontinuous across the faces of the underlying computational mesh.

	In this report, we devise  a higher order method for the numerical solution of the diffusion problem with discontinuous  coefficients on a general polyhedral mesh, which does not respect the position of multi-material interfaces.  The method builds upon the approximate static condensation (ASC) algorithm first proposed in~\cite{kikinzon2017approximate}; it admits high-contrast diffusion coefficients and discontinuous reconstruction of the interfaces as typical for the volume/moment-of-fluid method. The ASC approach uses the formulation of the interface diffusion problem as a system of local inter-cell problems  coupled through the concentration values and fluxes on the cell faces. This face concentration serves as the new primal unknown and it is further approximated by piecewise  polynomial functions defined on all faces of the global (unfitted polygonal) mesh. The original paper ~\cite{kikinzon2017approximate} considered $P_0$ (piecewise constant) approximation to the face concentration, while the present study introduces higher order polynomials for this purpose.
	Although the latter potentially leads to more accurate results  (and in some cases we shall see that it provides convergent results, when the $P_0$ version fails), it also raises some questions about stability and robustness, which we address in the present report.

	The approximate static condensation approach leaves much freedom in the choice of the discretization method for the local diffusion problems.  Posed in each polygonal cell, these problems are given by  diffusion equations with  discontinuous coefficients. The method benefits from the observation that   for any fixed cell $T$,  material interfaces, which possibly intersect with $T$,  induce a \emph{consistent polygonal subdivision} (local mesh) of $T$. Opposite to the global mesh, the local one is \emph{fitted} to interfaces, i.e. each micro polyhedral element is shared only by one material. Following~\cite{kikinzon2017approximate}, we apply the mimetic finite difference (MFD) method from~\cite{lipnikov2014mimetic} to approximate the local diffusion problems on this fine fitted mesh.
	After eliminating inter-cell degrees of freedom (d.o.f.), one obtains a system of linear algebraic equations for the face concentration degrees of freedom. The fluxes and inter-element degrees of freedom are recovered by simple postprocessing. In the report, we study the performance of the method in terms of accuracy, robustness with respect to the interfaces position,  and algebraic stability.

@@

	Further denote by~${{\vect \lambda}_\bcell \in \Rn{n_\bcell}}$ a vector of d.o.f. for elements of $\Lambda$ associated with all macro-faces spanning the boundary of~$\bcell$. We need a correspondence between elements of $\Lambda$ and local face elements $\vect \lambda_\mmesh$ used as boundary data in local problem \eqref{local}.
	This corresponding is given by the interpolation matrix $ \vect R_\mmesh$,
	\begin{equation}\label{asc_lambda}
		{\vect \lambda}_\mmesh = \vect R_\mmesh\,{\vect \lambda}_\bcell,
	\end{equation}
	where ${\vect \lambda}_\mmesh$ is effectively a piecewise constant approximation of ${\vect \lambda}_\bcell$ on $\mfaces_{\text{ext}}(T)$.
	Using this in \eqref{mini_flux_dofs}, we get
	\begin{equation}\label{asc_flux_lambda}
		{\vect u}^\text{ext}_\mmesh = \vect A_\mmesh\,\vect C_\mmesh\,\vect R_\mmesh\,{\vect \lambda}_\bcell + {\vect a}_\mmesh.
	\end{equation}
	%The interpolating matrix~$\vect R_\mmesh$ is completely defined by~$n$ and boundary mini\,faces~$\mfaces[ext]$ of~$\mmesh$.
	If we find the mortar vector ~${\vect \lambda}_\bcell$, we recover the local approximation to the face concentration, i.e. ${\vect \lambda}_\mmesh$, from~\eqref{asc_lambda}.
	Further we solve numerically local subproblems in~\eqref{dual_mini} applying the mimetic FD method. This completes the solution algorithm for \eqref{dual}.

>>>>

	Multi-phase and multi-material processes are ubiquitous in nature and engineering.
	Mathematical models of such processes are often formulated in terms of systems of partial differential equations with discontinuous coefficients and interface conditions on time-dependent internal boundaries.
	Due to their high complexity, these models are not suitable for analytical treatment, and hence numerical simulations became an indispensable tool for solving the models and gaining insights into underlying phenomena.

	Past decades evidenced an explosive growth in developing of efficient computational techniques for problems with evolving interfaces, which involve interaction between many materials or different phases of matter. This includes immersed boundary method \cite{peskin2002immersed}, level-set method \cite{sethian1999level}, and unfitted finite element methods \cite{belytschko2009review,burman2015cutfem}, to name several popular numerical approaches. However, numerous applications require local conservation of material volumes on computational meshes that do not explicitly track material interfaces. This is typical for simulations of high speed multi-material flows, where shocks and vorticity often require the use of the Arbitrary Lagrangian-Eulerian method. The complication with this method is that over the course of simulation, material interfaces become unaligned with the computational mesh, with their shape and position inside mesh cells not being explicitly known.

    The standard approach to dealing with cells intersected by material interfaces, also known as multi-material cells, is by using a homogenization of  material properties, where a single artificial material assigned to the cell has properties that are calculated based on a volumetric distribution of real materials contained in that cell. This allows to use standard discretization methods if one needs to solve for other physics, such as the heat transfer, on every time step of the simulation. Then, ad-hoc methods can be used to interpolate the solution for the contained materials based on the computed cell-average values~\cite{dawes2013solving}. This approach does not accurately model in-cell material interactions and can result in non-physical solutions when materials are substantially different.

    The alternatives generally require for the so-called interface reconstruction to be used in order to approximate material interfaces inside multi-material cells. Interface reconstruction methods typically represent materials in a cell by non-overlapping single-material polytopes that match given volume fractions. The most popular approach is based on using a single linear interface to represent an interface between two materials, and is commonly called the Piecewise-Linear Interface Calculation (PLIC). The volume-of-fluid (VOF)~\cite{hirt1981volume} is perhaps the most famous example of PLIC-type methods, which also include the moment-of-fluid (MOF) methods~\cite{ahn2009adaptive} that can provide better accuracy if centroids of materials are known. Recently introduced extended interface reconstruction methods, such as X-MOF~\cite{kikinzon2018xmof}, construct local meshes inside multi-material cells, where every cell of the local mesh contains only one material, and full local topology is available.

    The common drawback of interface reconstruction is that reconstructed material interfaces are discontinuous on the interfaces between multi-material cells. One approach to dealing with this issue is through merging the local meshes with the computational mesh, thus effectively refining the mesh so that every cell contains only one material~\cite{garimella2011supermesh}. This is a non-trivial computational geometry problem, and the resulting mesh in 2D typically contains very small and co-linerar faces, while the extension of this approach to 3D is infeasible for general unstructured meshes.

	In this article, we consider the numerical solution of the diffusion problem with discontinuous coefficients on polytopal meshes with multi-material cells. We assume that interface reconstruction is performed using an extended method, which provides local single-material meshes with full topological data for multi-material cells. One might consider solving this problem with Nitsche’s methods used in the framework of cut-cells~\cite{dolbow2012nitsche, burman2015cutfem, hansbo2002unfitted} or extended finite element methods (X-FEM)~\cite{dolbow2008xfem}. However, these methods are typically not locally conservative, and interface reconstruction is usually used when volume conservation is a requirement. Additionally, the discontinuity of material interfaces across faces of multi-material cells can be large enough to result in a significant loss of accuracy for these methods.

	We present a higher order method for solving this problem that builds upon the approximate static condensation (ASC) algorithm first proposed in~\cite{kikinzon2017approximate}; it admits high-contrast diffusion coefficients and discontinuous reconstruction of the interfaces. The ASC approach uses the formulation of the interface diffusion problem as a system of problems on local meshes that are coupled through the concentration values and fluxes on the faces of multi-material cells. This face concentration serves as the new primal unknown and it is further approximated by piecewise  polynomial functions defined on all faces of the global (unfitted polygonal) mesh. The original paper ~\cite{kikinzon2017approximate} considered $P_0$ (piecewise constant) approximation to the face concentration, while the present study introduces higher order polynomials for this purpose.
	Although the latter potentially leads to more accurate results  (and in some cases we shall see that it provides convergent results, when the $P_0$ version fails), it also raises some questions about stability and robustness, which we address in the present report.
	One might note that using $P_1$ (piecewise linear) approximation to the face concentration puts the proposed method into the framework of mortar methods~\cite{yotov2009mortar,mortar}. While those methods were developed for different types of problems and substantially differ in terms of implementation, but the proposed method fits their high-level formulation.

	The approximate static condensation approach leaves much freedom in the choice of the discretization method for the local diffusion problems.  Posed in each polygonal cell, these problems are given by diffusion equations with discontinuous coefficients. On the other hand, local meshes produces by extended interface reconstruction methods, unlike the the global mesh, contain only single-material cells. Following~\cite{kikinzon2017approximate}, we apply the mimetic finite difference (MFD) method from~\cite{lipnikov2014mimetic} to approximate the local diffusion problems on these local meshes.
	After eliminating inter-cell degrees of freedom (d.o.f.), one obtains a system of linear algebraic equations for the face concentration degrees of freedom. The fluxes and inter-element degrees of freedom are recovered by simple postprocessing. In the report, we study the performance of the method in terms of accuracy, robustness with respect to the interfaces position,  and algebraic stability.

@@

	Further denote by~${{\vect \lambda}_\bcell \in \Rn{n_\bcell}}$ a vector of d.o.f. for elements of $\Lambda$ associated with all macro-faces spanning the boundary of~$\bcell$. We need a correspondence between elements of $\Lambda$ and local face elements $\vect \lambda_\mmesh$ used as boundary data in local problem \eqref{local}.
	This corresponding is given by the interpolation matrix $ \vect R_\mmesh$,
	\begin{equation}\label{asc_lambda}
		{\vect \lambda}_\mmesh = \vect R_\mmesh\,{\vect \lambda}_\bcell,
	\end{equation}
	where ${\vect \lambda}_\mmesh$ is effectively a piecewise constant approximation of ${\vect \lambda}_\bcell$ on $\mfaces_{\text{ext}}(T)$.
	Using this in \eqref{mini_flux_dofs}, we get
	\begin{equation}\label{asc_flux_lambda}
		{\vect u}^\text{ext}_\mmesh = \vect A_\mmesh\,\vect C_\mmesh\,\vect R_\mmesh\,{\vect \lambda}_\bcell - {\vect a}_\mmesh.
	\end{equation}
	%The interpolating matrix~$\vect R_\mmesh$ is completely defined by~$n$ and boundary mini\,faces~$\mfaces[ext]$ of~$\mmesh$.
	If we find the mortar vector ~${\vect \lambda}_\bcell$, we recover the local approximation to the face concentration, i.e. ${\vect \lambda}_\mmesh$, from~\eqref{asc_lambda}.
	Further we solve numerically local subproblems in~\eqref{dual_mini} applying the mimetic FD method. This completes the solution algorithm for \eqref{dual}.
